A simplified dynamic bioenergetic model for coral-*Symbiodinium* symbioses

### 2016-09-20 JTB_ms/JTB_ms.Rmd
move to subdir

bibliography: ~/library.bib
output: 
  rticles::elsevier_article:
    fig_caption: yes
header-includes:
  - \usepackage{tabularx}
---

@@

```{r Fig1, fig.width=5, fig.height=5, echo=FALSE, fig.cap="Graphical representation of coral-algal symbiosis model."}
library(png)
library(grid)
img <- readPNG("../img/model.png")
grid.raster(img)
```

@@

where $j_{HGm}$ is the maximum specific growth rate. Finally, the rate of change in coral biomass is calculated as the specific growth rate minum the specific biomass turnover rate (i.e., maintenance costs), according to:

\begin{equation} {dH \over dt} = j_{HG} - j_{HT}^0 \end{equation}

@@

\begin{equation} \rho_N = j_N + n_{NX}j_X + r_{NH} - n_{NH}j_{HG} \end{equation}

@@

\begin{equation} j_{CP} = \bigg({1 \over j_{CPm}} + {1 \over {n_{LC} j_L }} + {1 \over {(j_{CO_2} + r_{CH}){H \over S} + r_{CS}}} - {1 \over {n_{LC} j_L + (j_{CO_2} + r_{CH}){H \over S} + r_{CS}}}\bigg)^{-1} \div c_{ROS} \end{equation}

@@

\begin{equation} j_{eL} = j_L - j_{CP} / n_{LC} \end{equation}

@@

\begin{equation} \rho_C = j_{CP} - j_{SG} \end{equation}

@@

This second component of symbiont biomass loss can represent both photodamage and/or symbiont expulsion (i.e., bleaching), both of which happen in response to higher levels or ROS production. The constant 5 is included to increase biomass loss in response to ROS. (Note that recycling of symbiont biomass turnover ($rNS$ and $rCS$) only occurs based on the basal maintenance related turnover (i.e., $j_{ST}^0$), and not the bleaching-related biomass loss, as this loss represents biomass being damaged or being expelled from the holobiont). Finally, the balance equation representing the change in symbiont biomass over time can be specified as the growth rate minus the turnover rate:

\begin{equation} {dS \over dt} = j_{SG} - j_{ST} \end{equation}

@@

The differential equations specified in 7 and 14 were solved using an implicit time-stepping method because of the implicit variables of surplus sharing. Euler method?

@@

>>>>

bibliography: library.bib
output: 
  rticles::elsevier_article:
    fig_caption: yes
---
```{r setup, include=FALSE}
#knitr::opts_knit$set(root.dir = normalizePath(".."))
```

```{r load, include=FALSE}
# Source default parameters and functions
sapply(list.files(path="../../R", pattern="*.R", full.names=T), source, .GlobalEnv)
```

@@

![Graphical representation of coral-algal symbiosis model.](../img/model.png)

@@

@@

\begin{equation} \rho_N = (j_N + n_{NX}j_X + r_{NH} - n_{NH}j_{HG})_+ \end{equation}

@@

\begin{equation} j_{CP} = \bigg({1 \over j_{CPm}} + {1 \over {n_{LC} j_L }} + {1 \over {(j_{CO_2} + r_{CH}){H \over S} + r_{CS}}} - {1 \over {n_{LC} j_L + (j_{CO_2} + r_{CH}){H \over S} + r_{CS}}}\bigg)^{-1} \cdot c_{ROS}^{-1} \end{equation}

@@

\begin{equation} j_{eL} = (j_L - j_{CP} / n_{LC})_+ \end{equation}

@@

\begin{equation} \rho_C = (j_{CP} - j_{SG})_+ \end{equation}

@@

This second component of symbiont biomass loss can represent both photodamage and/or symbiont expulsion (i.e., bleaching), both of which happen in response to higher levels or ROS production. The constant 5 is included to increase biomass loss in response to ROS. (Note that recycling of symbiont biomass turnover ($rNS$ and $rCS$) only occurs based on the basal maintenance related turnover (i.e., $j_{ST}^0$), and not the bleaching-related biomass loss, as this loss represents biomass being damaged or being expelled from the holobiont).

## Model state equations
Finally, the balance equations representing the specific growth rates of symbiont and host biomass over time can be expressed as:

\begin{equation} {dS \over Sdt} = j_{SG} - j_{ST} \end{equation}

\begin{equation} {dH \over Hdt} = j_{HG} - j_{HT}^0 \end{equation}

@@

A time-stepping Euler method was used to solve the state equations since the production and rejection fluxes of the SUs are imiplicitly defined. Specifically, the rejection fluxes of carbon and nitrogen from the symbiont and host biomass SUs act as reciprocal input fluxes to the other SU. In addition, the rejection flux of excitation energy from the photosynthesis SU acts to reduce its own production flux (i.e., photoinhibition), and hence a discretized time-stepping procedure was necessary. A vector of time values was created for each simulation run, along which dynamic environmental forcing functions (irradiance, DIN, and prey abundance) can be designed. These vectors, along with initial values of symbiont and host biomass, then serve as input to the time-stepping function, which solves for the current system state using values of the previous system state where necessary. A default time step of 0.1 days was used for all simulations, which were performed using R code that is available in the data repository accompanying this article: [github.com/jrcunning/Rcoral](github.com/jrcunning/Rcoral).

@@

To analyze qualitative behavior, we ran the model to steady state across gradients of external irradiance and DIN, and plotted the specific growth rate (Fig. 2A) and symbiont to host biomass ratio (Fig. 2B). Results are consistent with observations in corals: specific growth rates

![Steady state values of (A) specific growth ($Cmol \cdot Cmol^{-1} \cdot d^{-1}$) and (B) the symbiont to host biomass ratio ($CmolS \cdot CmolH^{-1}$) across gradients of external irradiance and dissolved inorganic nitrogen. Simulations for each combination of light and nutrients (21 points along each axis) were run for 100 days with a time step of 1 day. Negative steady state growth rates, and corresponding S:H ratios, were set to zero.](../img/Fig2.png)
